---
title: "Data Wrangling"
author: "Student name submitting the assessment report comes here"
subtitle: Practical assessment 2
date: ""
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---


## **Setup**

```{r}

# Load the necessary packages required to reproduce the report. For example:

library(dplyr)
library(kableExtra)
library(magrittr)
library(readr)


```


## **Student names, numbers and percentage of contributions**
```{r, echo=FALSE}

# Add your names, numbers and percentage of your contribution here.

na<- c(" Karim Aly"," Bhanuchandu Bochu")
no<- c(" 4104031","  4120371")
pc<- c("Percentage1","  Percentage2")

s<- data.frame(cbind(na,no,pc))
colnames(s)<- c("Student name", "Student number", "Percentage of contribution")

s %>% kbl(caption = "Group information") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```
<br>
<br>

## **Executive Summary**

In your own words, provide a brief summary of the preprocessing. Explain the steps that you have taken to preprocess your data. Write this section last after you have performed all data preprocessing. (Word count Max: 300 words)
You may use bulleted lists like this:

* a
* b
* ....

<br>
<br>

## **Data**

Provide explanations here.

```{r}

# Import the data, provide your R codes here.

# Import the data without column names
us_gdp_by_state_df <- read_csv("us-gdp-by-state.csv", show_col_types = FALSE)

#aus_real_estate_df$City <- toupper(aus_real_estate_df$City)


# Display the first few rows of the data frame
print(head(us_gdp_by_state_df))



```

```{r}
us_realtor_df <- read_csv("us-realtor-data.csv", show_col_types = FALSE)



print(head(us_realtor_df))

```

```{r}
merged_df = merged_df <- left_join(us_realtor_df, us_gdp_by_state_df, by = c("state" = "Area"))



print(tail(merged_df))
```

<br>
<br>

## **Understand** 

```{r}
print(str(merged_df))
```

```{r}
merged_df$status <- factor(merged_df$status, 
                      levels = c(1, 2, 3), 
                      labels = c("for_sale", "sold", "ready_to_build"))


merged_df$prev_sold_date <- as.Date(merged_df$prev_sold_date, format = "%d-%m-%Y")

print(str(merged_df))

```

```{r}
# Filter the dataset to include only Alabama
alabama_df <- merged_df[merged_df$state == "Alabama", ]

# Display the first few rows of the Alabama dataset
print(head(alabama_df))

# Display the dimensions of the Alabama dataset
print(dim(alabama_df))
```

```{r}


# This is the R chunk for the Understand Section

```

Provide explanations here. 

<br>
<br>

##	**Tidy & Manipulate Data I **


```{r}

# This is the R chunk for the Tidy & Manipulate Data I 

```

Provide explanations here. 

<br>
<br>

## **Tidy & Manipulate Data II** 

```{r}


###Question NO 6


# Create a new variable 'price_per_sqft' using base R
merged_df$price_per_sqft <- merged_df$price / merged_df$house_size

# Display the first few rows of the updated dataframe
print(head(merged_df))

# Summary statistics of the new variable
summary(merged_df$price_per_sqft)

print(tail(merged_df))

# This is the R chunk for the Tidy & Manipulate Data II 

```

Provide explanations here. 

<br>
<br>

##	**Scan I **
```{r}
# Filter the dataset to include only Alabama
alabama_df <- merged_df[merged_df$state == "Alabama", ]

# Display the first few rows of the Alabama dataset
print(head(alabama_df))

# Display the dimensions of the Alabama dataset
print(dim(alabama_df))
```

```{r}

# This is the R chunk for the Scan I

```

Provide explanations here. 

<br>
<br>

##	**Scan II**

```{r}

# This is the R chunk for the Scan II

```

Provide explanations here. 

<br>
<br>

##	**Transform **

```{r}

# This is the R chunk for the Transform Section

```

Provide explanations here. 


<br>
<br>

##	**Presentation **

Add the link to your presentation inside the brackets below.

[Presentation](...)

