---
title: "Data Wrangling"
author: "Student name submitting the assessment report comes here"
subtitle: Practical assessment 2
date: ""
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---


## **Setup**

```{r}

# Load the necessary packages required to reproduce the report. For example:

library(kableExtra)
library(magrittr)
library(readr)


```


## **Student names, numbers and percentage of contributions**
```{r, echo=FALSE}

# Add your names, numbers and percentage of your contribution here.

na<- c(" Name1"," Name2")
no<- c(" number1","  number2")
pc<- c("Percentage1","  Percentage2")

s<- data.frame(cbind(na,no,pc))
colnames(s)<- c("Student name", "Student number", "Percentage of contribution")

s %>% kbl(caption = "Group information") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```
<br>
<br>

## **Executive Summary**

In your own words, provide a brief summary of the preprocessing. Explain the steps that you have taken to preprocess your data. Write this section last after you have performed all data preprocessing. (Word count Max: 300 words)
You may use bulleted lists like this:

* a
* b
* ....

<br>
<br>

## **Data**

Provide explanations here.

```{r}

# Import the data, provide your R codes here.

# Import the data without column names
adult_df <- read_csv("adult/adult.data", show_col_types = FALSE, col_names = FALSE)

# Assign column names directly
colnames(adult_df) <- c("age", "workclass", "fnlwgt", "education", "education-num", 
                        "marital-status", "occupation", "relationship", "race", 
                        "sex", "capital-gain", "capital-loss", "hours-per-week", 
                        "native-country", "income")

# Display the first few rows of the data frame
print(head(adult_df))



```
```{r}
census_df <- read_csv("census/census-income.data", show_col_types = FALSE)

colnames(census_df) <- c(
  "age", "class_of_worker", "industry_code", "occupation_code", "adjusted_gross_income",
  "education", "wage_per_hour", "enrolled_in_edu_inst_last_wk", "marital_status", "major_industry_code",
  "major_occupation_code", "mace", "hispanic_Origin", "sex", "member_of_a_labor_union",
  "reason_for_unemployment", "full_or_part_time_employment_stat", "capital_gains", "capital_losses",
  "dividends_from_stocks", "federal_income_tax_liability", "tax_filer_status",
  "region_of_previous_residence", "state_of_previous_residence", "detailed_household_and_family_stat",
  "detailed_household_summary_in_household", "instance_weight", "migration_code_change_in_msa",
  "migration_code_change_in_reg", "migration_code_move_within_reg", "live_in_this_house_1_year_ago",
  "migration_prev_res_in_sunbelt", "num_persons_worked_for_employer", "family_members_under_18",
  "total_person_earnings", "country_of_birth_father", "country_of_birth_mother", "country_of_birth_self",
  "citizenship", "total_person_income", "own_business_or_self_employed", "taxable_income_amount",
  "fill_inc_questionnaire_for_veterans_admin", "veterans_benefits", "weeks_worked_in_year"
)

print(head(census_df))

#install.packages("dplyr")

# load the dplyr package

library(dplyr)

census_df <- select(census_df, 
  -tax_filer_status, -dividends_from_stocks, -capital_losses, -capital_gains, 
  -migration_code_change_in_reg, -migration_code_change_in_msa, -migration_code_move_within_reg, 
  -migration_prev_res_in_sunbelt, -fill_inc_questionnaire_for_veterans_admin, 
  -veterans_benefits, -weeks_worked_in_year
)

print(head(census_df))
```

<br>
<br>

## **Understand** 


```{r}

# This is the R chunk for the Understand Section

```

Provide explanations here. 

<br>
<br>

##	**Tidy & Manipulate Data I **


```{r}

# This is the R chunk for the Tidy & Manipulate Data I 

```

Provide explanations here. 

<br>
<br>

## **Tidy & Manipulate Data II** 

```{r}

# This is the R chunk for the Tidy & Manipulate Data II 

```

Provide explanations here. 

<br>
<br>

##	**Scan I **

```{r}

# This is the R chunk for the Scan I

```

Provide explanations here. 

<br>
<br>

##	**Scan II**

```{r}

# This is the R chunk for the Scan II

```

Provide explanations here. 

<br>
<br>

##	**Transform **

```{r}

# This is the R chunk for the Transform Section

```

Provide explanations here. 


<br>
<br>

##	**Presentation **

Add the link to your presentation inside the brackets below.

[Presentation](...)

